---
title: "Reproducible Research: Peer Assessment 1"
subtitle: "Subtitle"
author: "Daniel F Seg"
output: 
  html_document:
    keep_md: true
---

### **Libraries**

Those libraries were used to work with the dataset.

```{r libraries, echo = TRUE}
library(tidyverse)
```

### Loading and preprocessing the data

This process considers that the dataset is unzipped on data folder on your working directory.

```{r loading, echo=TRUE}
activity <- read_csv("data/activity.csv", 
                     na = "NA",
                     col_types = cols(steps = col_integer(), 
                                      date = col_date(format = "%Y-%m-%d"), 
                                      interval = col_integer()))
```

### Preprocessing the data

```{r processing, echo=TRUE}

summary(activity)

```

### What is mean total number of steps taken per day?

```{r mean, echo=TRUE}

act_mean <- activity %>%
    
    group_by(date) %>%

    summarise(sum = sum(steps, na.rm = TRUE), 
              mean = mean(steps, na.rm = TRUE)) %>%
    
    filter(sum != 0)
    
daily_mean <- act_mean %>% summarise(mean = mean(mean, na.rm = TRUE)) %>% as.numeric()

act_mean %>%    
        
    ggplot() +
    
    geom_line(aes(date, mean))



```

Daily `r daily_mean`

### What is the average daily activity pattern?



### Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
