---
title: "R Mardown with step counting data"
subtitle: "Reproducible Research: Peer Assessment 1 - week 2"
author: "Daniel F Seg"
output: 
  html_document:
    keep_md: true
---

```{r locale, echo = FALSE, results = 'hide'}

# Setting localization

Sys.setlocale(category = "LC_ALL", locale = "English_United Kingdom")

```

### **Libraries**

Those libraries has been used to work with the dataset.

```{r libraries, echo = TRUE, results = 'hide'}

library(tidyverse)

```

### **Loading and preprocessing the data**

This process considers that the dataset is unzipped onto `\data` folder on your working directory. I'm also adding two new variables, `weekday` and `is.weekend`, to categorize the database later on.

```{r loading, echo = TRUE}

activity <- read_csv("data/activity.csv", 
                     na = "NA",
                     col_types = cols(steps = col_integer(), 
                                      date = col_date(format = "%Y-%m-%d"), 
                                      interval = col_integer())) %>%
  
  mutate(weekday = as.factor(weekdays(date))) %>%
  
  mutate(is.weekend = if_else(weekday %in% c("Saturday", "Sunday"), 
                              "weekend", 
                              "weekday") %>% as.factor())

summary(activity)

```

```{r, echo = FALSE, results = 'hide'}

na_tot <- sum(is.na(activity$steps))

```

Summarizing the data set, it is possible to notice a total of `r na_tot` `NA`s on the steps variable.


### **What is mean total number of steps taken per day?**

For this topic a new data frame has been created

```{r mean, echo=TRUE}

act_mean <- activity %>%
    
    group_by(date, weekday, is.weekend) %>%

    summarise(sum = sum(steps, na.rm = TRUE), 
              mean = mean(steps, na.rm = TRUE))
   

act_mean %>%
  
  na.omit() %>%
  
  ggplot() +
  
  geom_histogram(aes(sum), bins = 15, fill = "lightgreen", color = "black") + 
  
  geom_hline(yintercept = 0, size = 1.5) +
  
  geom_vline(xintercept = mean_steps <- mean(act_mean$sum), color = "red", size = 1) +

  geom_vline(xintercept = median_steps <- median(act_mean$sum), color = "purple", size = 1) + 
  
  labs(title = "Histogram of steps per day",
       subtitle = "Median (purple), Mean (red)",
       caption = "by DFS") +
  
  xlab("Steps per day") +
  
  ylab("Frequency") +
  
  theme_bw()
  
```
  
The daily mean and median steps are **`r mean_steps`** and **`r median_steps`**, respectively.
  
### **What is the average daily activity pattern?**  
  
  I have created the `daily pattern` Dataframa using summarise and mean().  
  

```{r pattern, echo = TRUE}


daily_pattern <- activity %>%    # Daily steps pattern
  
  group_by(interval) %>%
  
  summarise(mean_steps = mean(steps, na.rm = TRUE))

summary(daily_pattern)

```

  And plotted the data:

```{r daily, echo = TRUE}

daily_pattern %>%
  
  na.omit() %>%
  
  ggplot() +
  
  scale_x_continuous(expand = c(0,0), n.breaks = 10) +

  coord_cartesian(xlim = c(50, 2355)) + 
  
  geom_line(aes(interval, mean_steps), size = .8, color = "blue") + 
  
  labs(title = "Number of steps by 5-min interval",
     subtitle = "Mean of all intervals along a day",
     caption = "by DFS") +
  
  xlab("5-min data collection interval") +
  
  ylab("Number of steps") +
  
  theme_bw()

```


The following code has been used to extract the largest step count by interval:

```{r max, echo = TRUE}
max_steps_interval <- daily_pattern %>% filter(mean_steps == max(mean_steps))
```

The interval which has occurred the largest amount of steps is **`r max_steps_interval$interval`** with **`r round(max_steps_interval$mean_steps)`** steps counted.


### **Imputing missing values**

There are a total of `r na_tot` `NA`s on the step values. The imputation method is replacing NA for the mean value for all intervals.

```{r impute_na, echo = TRUE}

na_imputed <- activity %>%
  
  filter(is.na(steps)) %>%
  
  mutate(steps = as.integer(mean(activity$steps, na.rm = TRUE))) %>%
  
  rbind(., activity %>% filter(!is.na(steps)))

na_imputed %>%
  
 group_by(date, weekday, is.weekend) %>%

    summarise(sum = sum(steps, na.rm = TRUE), 
              mean = mean(steps, na.rm = TRUE)) %>%
  
  ggplot() +
  
  geom_histogram(aes(sum), bins = 15, fill = "lightblue", color = "black") + 
  
  geom_hline(yintercept = 0, size = 1.5) +
  
  labs(title = "Histogram of steps per day",
       subtitle = "Imputed data",
       caption = "by DFS") +
  
  xlab("Steps per day") +
  
  ylab("Frequency") +
  
  theme_bw()

```

### **Are there differences in activity patterns between weekdays and weekends?**

Using the dataframe `na_imputed`, week/weekend have been filtered and plot to show the different patterns.

```{r weekends_plot, echo = TRUE}

na_imputed %>%
  
  group_by(is.weekend, interval) %>%
  
  summarise(means = mean(steps)) %>%
  
  ggplot() +
  
  scale_x_continuous(expand = c(0,0), n.breaks = 10) +

  geom_line(aes(interval, means), size = .8, color = "blue") + 
  
  facet_grid(is.weekend ~ .) +
  
  labs(title = "Steps patterns weekdays / weekends",
     subtitle = "NA imputed",
     caption = "by DFS") +
  
  xlab("5-min data collection interval") +
  
  ylab("Number of steps") +
  
  theme_bw()




```



